---
title: "MThambeliyagodage_Data605_W11_Assign11"
author: "Matheesha Thambeliyagodage"
date: "April 22, 2018"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document: 
    toc: yes
  word_document: 
    toc: yes
  html_notebook: 
    toc: yes
---
```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Data 605 Assignment 11

## Q1
Using the `cars` dataset in R, build a linear model for stopping distance as a function of speed and
replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and
residual analysis).
```{r}
data(cars)
head(cars)
summary(cars)
colnames(cars)
ncol(cars)
nrow(cars)
```

## Data Visualization:
```{r}
cars_speed_df = arrange(cars, speed)

ggplot(data=cars_speed_df, aes(cars_speed_df$speed)) + 
  geom_histogram(aes(fill = ..count..)) +
  scale_fill_gradient("Count", low = "blue", high = "pink") +
  labs(title = "Historgram - Speed") +
  labs(x = "speed") +
  labs(y = "Count")
```

```{r}
cars_dist_df = arrange(cars, dist)

ggplot(data=cars_dist_df, aes(cars_dist_df$dist)) + 
  geom_histogram(aes(fill = ..count..)) +
  scale_fill_gradient("Count", low = "blue", high = "purple") +
  labs(title = "Historgram - Distance") +
  labs(x = "dist") +
  labs(y = "Count")
```
## Statistical Analysis:

### Plot of two variables Speed & distance
The Speed is **explanatory** variable and stopping distance is the **response**.
```{r}
plot(cars$speed, cars$dist, xlab='Speed (mph)', ylab='Stopping Distance (ft)', 
     main='Stopping Distance vs. Speed')
```

### Display the Linear Model

```{r}
cars_lm <- lm(cars$dist ~ cars$speed)
cars_lm

plot(cars$speed, cars$dist, xlab='Speed (mph)', ylab='Stopping Distance (ft)', 
     main='Stopping Distance vs. Speed')
abline(cars_lm)
```

The above appears some correlation between two variables, but let us evaluate the linear model we have. 

```{r}
summary(cars_lm)
```

The median value of the residuals is somewhat close to zero and quartiles and min/max values are roughly the same magnitude. The standard error of the `speed` variable is more than 9 times smaller than the corresponding coefficient. There should not be a lot of variability in this coefficient. On the other hand, the difference between the intercept estimate and standard error is less significant, so there may be more variability. The `speed` coefficient is highly significant. The intercept coefficient is less significant, but it is still relevant depending on the confidence interval desired. Finally, $R^2$ explains about 65.11% of the data's variation.

```{r}
plot(cars_lm$fitted.values, cars_lm$residuals, xlab='Fitted Values', ylab='Residuals')
abline(0,0)
```

It is possible to say that the outlier values do not show the same variance of the residuals; however, it is not very clear. I think it is reasonable to continue with the analysis and assume similar variance of residuals.

```{r}
qqnorm(cars_lm$residuals)
qqline(cars_lm$residuals)
```

Althought again there are some problems at the outlier levels, the normal Q-Q plot of the residuals appears to follow the theoretical line. Residuals are reasonably normally distributed.

### Conclusion

I believe the linear model does a good job at explaining the data. There appears to be some slight curvature in the main plot and in the residuals plot, so I decided to try a simple quadratic model (see below). It has it's own problems - again varability of residuals is not constant enough, q-q plot has some deviations, coefficients are not very significant and $R^2$ is not increased by much. I don't think it's an improvement over a simplier linear model.

#### Quadtratic Model

```{r}
speed <- cars$speed
speed2 <- speed^2
dist <- cars$dist
  
cars_qm <- lm(dist ~ speed + speed2)
summary(cars_qm)

speedvalues <- seq(0, 25, 0.1)
predictedcounts <- predict(cars_qm,list(speed=speedvalues, speed2=speedvalues^2))

plot(speed, dist, pch=16, xlab='Speed (mph)', ylab='Stopping Distance (ft)')
lines(speedvalues, predictedcounts)

plot(cars_qm$fitted.values, cars_qm$residuals, xlab='Fitted Values', ylab='Residuals')
abline(0,0)

qqnorm(cars_qm$residuals)
qqline(cars_qm$residuals)
```