---
title: "DATA606_W1_Lab_1_MatheeshaThambeliyagodage"
author: "Matheesha Thambeliyagodage"
date: "February 9, 2017"
output: html_document
---
Begin by loading the data set of 20,000 observations into the R workspace.
```{r}
library(DATA606)
#source("more/cdc.R")
#source("https://github.com/mathsanu/CUNY_MSDA/blob/master/DATA606/W1/lab_1/cdc.R")
source("http://www.openintro.org/stat/data/cdc.R")
names(cdc)
head(cdc)
```
1.Make a scatterplot of weight versus desired weight. Describe the relationship between these two variables.
```{r}
plot(cdc$weight, main="Scatter Plot -  Weight versus Desired Weight of Passengers", cdc$wtdesire, xlab = "Weight", ylab = "Desired Weight",  col="purple", las = 1)
```
It appears to be a general positive correlation between the two variables weight versus desired weight, when desired weight increases weight also increses,although there is a great deal of spread. I assume they could be due to data entry mistakes (eg 700lb).
--
2. Let's consider a new variable: the difference between desired weight (wtdesire) and current weight (weight). Create this new variable by subtracting the two columns in the data frame and assigning them to a new object called wdiff.
```{r}
cdc$wdiff <- cdc$wtdesire - cdc$weight
#cdc$wdiff
names(cdc)
head(cdc)
```

3.What type of data is wdiff? If an observation wdiff is 0, what does this mean about the person's weight and desired weight. What if wdiff is positive or negative?

wdiff is a scale variable an it is integer discrete because weights are integer values. Each record hase a wdiff value.

wdiff is equal to 0 means the person has the oplimal weight to the height or in their desired weight. wdiff > 0
then their weight is below their desired value, while wdiff < 0 implies that their actual weight is above their
desired value.
--------------------
4. Describe the distribution of wdiff in terms of its center, shape, and spread, including any plots you use. What does this tell us about how people feel about their current weight?

```{r}
summary(cdc$wdiff)
```
```{r}
table(cdc$wdiff)
```
```{r}
hist(cdc$wdiff, ,main="Histogram for wdiff", breaks = 100,xlab="Gap between Desired Weight  and Actual Weight ", ylab = "Person Count",border="blue",     col="yellow")
```
--This distribution shows the "Gap between Desired  Weight and Actual Weight" heavily distributed between 0-10. And majority of the people are overweight. Median is -10 with the Mean -14.5

--5.Using numerical summaries and a side-by-side box plot, determine if men tend to view their weight differently than women.

```{r}
summary(cdc$wdiff[cdc$gender == "m"])
```

```{r}
summary(cdc$wdiff[cdc$gender == "f"])
```

```{r}
boxplot(main="Box Plot for Men and Wemen weight Distribution", cdc$wdiff ~ cdc$gender, xlab="Gender",  ylab = "wdiff",col="Red", outline = F)
```
Above box plot and summary calculations shows median of women is lower than men. This means women are away from the desired weight than men (overweight women)

--6.Now it's time to get creative. Find the mean and standard deviation of weight and determine what proportion of the weights are within one standard deviation of the mean.

```{r}
mean_weight <- mean(cdc$weight)
mean_weight
```

```{r}
sd_weight <- sd(cdc$weight)
sd_weight
```
```{r}
weights_within_one_sd <- subset(cdc, weight < mean_weight + sd_weight & weight > mean_weight -
+ sd_weight)
weights_within_one_sd
```
```{r}
dim(weights_within_one_sd)
```
There are 14152 observations in the required subset, out of a total of 20000 observations, this represents a
proportion of 14152/20000 = 0.7076, or about 71%. Thus 71% of weights are within 1 standard deviation of the
mean.
