---
title: "MThambeliyagodage_Data605_W3_Assign3"
author: "Matheesha Thambeliyagodage"
date: "February 18, 2018"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

##  <span style="color:Green"> Problem set 1 </span>
### Q1
#### <span style="color:Green">What is the rank of the matrix A?</span>

$$ \begin{aligned}
           \begin{bmatrix} 1 & 2 & 3 & 4 \\
                        -1 & 0 & 1 & 3 \\
                        0 & 1 & -2 & 1 \\
                        5 & 4 & -2 & -3
        \end{bmatrix} 
    \end{aligned}
$$
```{r}
require(Matrix)
A <-  matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), ncol = 4,byrow = T)
rankMatrix(A)
```
#### <span style="color:Red">The the rank of the matrix A is 4 </span>
### Q2
#### <span style="color:Green">Given an mxn matrix where m > n, what can be the maximum rank? The mini-
mum rank, assuming that the matrix is non-zero?</span>

* let A be an mxn matrix.  Then the row rank of A is equal
to the maximum number of linear independent row vectors of A. 
Hence the row-rank of mxn matrix is at most m.

* Similarly, the column-rank of A is the maximum number of linear 
independent columns of A. Hence the column-rank of A is at most n.

* Since for any matrix A mxn over field F, the row-rank = column-rank. Therefore
to be the common value of row-rank A and column-rank A, clearly rank A has to be 
equal or less of  Min (m,n) for any mxn matrix

* From the above, for any m x n matrix, if m is greater than n, then the maximum rank of the matrix is m.
and if m is less than n, then the maximum rank of the matrix is n.

* The rank of a matrix would be zero only if the matrix had no elements. 
However, If a matrix had even one element, its minimum rank would be one.

### Q3
#### <span style="color:Green">What is the rank of matrix B?</span>
$$ \begin{aligned}
           \begin{bmatrix} 1 & 2 & 1 \\
                        3 & 6 & 3 \\
                        2 & 4 & 1 
            \end{bmatrix} 
    \end{aligned}
$$
```{r}
B <-  matrix(c(1, 2, 1, 3, 6, 3, 2, 4, 2), ncol = 3, byrow = T)
rankMatrix(B)
```
#### <span style="color:Red">The the rank of the matrix B is 1 </span>
## Problem set 2
* Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your
work. You'll need to write out the characteristic polynomial and show your solution.

$$ A = \begin{aligned}
           \begin{bmatrix} 1 & 2 & 3 \\
                        0 & 4 & 5 \\
                        0 & 0 & 6 
            \end{bmatrix} 
    \end{aligned}
$$
$$ \text{ Definition: A scalar} \lambda  \text{ is called an eigenvalue of the n × n matrix A is there is a nontrivial solution
x of Ax =} \lambda x \\ \text{ . 
Such an x is called an eigenvector corresponding to the eigenvalue} \lambda.\\
\text{ A scalar} \lambda \text{ is an eigenvalue of an n×n matrix A if and only if} \lambda \text{ satisfies the characteristic equation.}\\
 det( A-\lambda I) = 0
$$ 
$$
\lambda I = \begin{aligned}
           \begin{bmatrix} \lambda & 0 & 0 \\
                        0 & \lambda & 0 \\
                        0 & 0 & \lambda 
            \end{bmatrix} 
\end{aligned}
$$
$$
A-\lambda I = \begin{aligned}
          \begin{bmatrix} 1 & 2 & 3 \\
                        0 & 4 & 5 \\
                        0 & 0 & 6 
            \end{bmatrix} 
            -
           \begin{bmatrix} \lambda & 0 & 0 \\
                        0 & \lambda & 0 \\
                        0 & 0 & \lambda 
            \end{bmatrix} 
            =
            \begin{bmatrix} 1-\lambda & 2 & 3 \\
                        0 & 4-\lambda & 5 \\
                        0 & 0 & 6 -\lambda 
            \end{bmatrix} 
            = 0
\end{aligned}
$$
* We can calculate the characteristic polymomial to be:

$$
\begin{aligned} Det(A-\lambda I) &=  (\lambda -1)((\lambda -4)(\lambda -6) -0) + 2((0-0) -3(0-0)) = 0\\
               & = \lambda^3 - 11\lambda^2 + 34\lambda - 24 \\ &= (\lambda -1)(\lambda -4)(\lambda-6)\\
               & = 0 \end{aligned}
$$
$$ \text{Setting this equal to 0 and solving for} \lambda  \text{ we get that } \lambda \text{ = 1, 4,6.  
These are the three eigenvalues of A.}
$$

#### <span style="color:Green"> Calculate the eigenvectors of A using $\lambda$  </span>

Now, lets plug in each value of $\lambda$ into the matrix:

<span style="color:Green"> $\lambda = 1$ </span>
------------

$( \lambda I_3 - A ) = 
\left[ {\begin{array}{ccc}
0 & -2 & -3\\
0 & -3 & -5\\
0 & 0 & -5\\
\end{array} } \right]$

```{r}

A1 <- matrix(c(0,0,0,-2,-3,0,-3,-5,-5),nrow=3)

# eliminate A1[2,3] by subtracting row 3 from row 2
# eliminate A1[1,3] by subtracting 3/5 * row 3 from row 

A1[2,] <- A1[2,] - A1[3,]

A1[1,] <- A1[1,] - (3/5)*A1[3,]

A1

```

```{r}
# eliminate A1[1,2] by subtracting (2/3)*A1[2,]

A1[1,] <- A1[1,] - (2/3)*A1[2,]

A1

```

```{r}

# Get matrix into reduced row echelon form by making pivots = 1
# Then move the zero rows to the bottom

A1[2,] <- A1[2,]/-3
A1[3,] <- A1[3,]/-5

A1 <- A1[c(2,3,1),]

```

$( \lambda I_3 - A )v = 
\left[ {\begin{array}{ccc}
0 & 1 & 0\\
0 & 0 & 1\\
0 & 0 & 0\\
\end{array} } \right]
\left[ {\begin{array}{c}
v_1\\
v_2\\
v_3\\
\end{array} } \right]$

$v_2 = 0 \\ v_3 = 0$

Let $v_1 = t$

$E_{\lambda = 1} = t \left[ {\begin{array}{c}
1\\
0\\
0\\
\end{array} } \right]$

where $t$ is real

***

<span style="color:Green"> $\lambda = 4$ </span>
------------


$( \lambda I_3 - A ) = 
\left[ {\begin{array}{ccc}
3 & -2 & -3\\
0 & 0 & -5\\
0 & 0 & -2\\
\end{array} } \right]$

```{r}

A4 <- matrix(c(3,0,0,-2,0,0,-3,-5,-2),nrow=3)

# Eliminate A4[2,3] by subtracting (5/2) * row 3 from row 2
# Eliminate A4[1,3] by subtracting (3/2) * row 3 from row 1

A4[2,] <- A4[2,] - (5/2)*A4[3,]
A4[1,] <- A4[1,] - (3/2)*A4[3,]

A4

```

```{r}

# Divide row 1 by 3 to make the pivot 1
# Divide row 3 by -2 to make the pivot 1
# Move the zero row to the bottom

A4[1,] <- A4[1,]/3
A4[3,] <- A4[3,]/-2

A4 <- A4[c(1,3,2),]


```

$( \lambda I_3 - A )v = 
\left[ {\begin{array}{ccc}
1 & -\frac{2}{3} & 0\\
0 & 0 & 1\\
0 & 0 & 0\\
\end{array} } \right]
\left[ {\begin{array}{c}
v_1\\
v_2\\
v_3\\
\end{array} } \right]$

$v_1 -\frac{2}{3}v_2 = 0 \\ v_3 = 0$

let $v_2 = t$

So:
$\frac{2}{3}t = v_1\\$

$E_{\lambda = 4} = t \left[ {\begin{array}{c}
\frac{2}{3}\\
1\\
0\\
\end{array} } \right]$

Where t is real

***

<span style="color:Green"> $\lambda = 6$ </span>
------------


```{r}

A6 <- matrix(c(5,0,0,-2,2,0,-3,-5,0),nrow=3)

# Eliminate A6[1,2] by subtracting -1 * row 2

A6[1,] <- A6[1,] + A6[2,]

# Divide Row 1 by 5 and row 2 by 2 to make the pivots 1

A6[1,] <- A6[1,]/5
A6[2,] <- A6[2,]/2

A6

```

$( \lambda I_3 - A )v = 
\left[ {\begin{array}{ccc}
1 & 0 & -1.6\\
0 & 1 & -2.5\\
0 & 0 & 0\\
\end{array} } \right]
\left[ {\begin{array}{c}
v_1\\
v_2\\
v_3\\
\end{array} } \right]$

$v_1 - 1.6v_3 = 0 \\ v_2 - 2.5v_3 = 0$

let $v_3 = t$

So

$V_1 = 1.6t \\ v_2 = 2.5t$

$E_{\lambda = 6} = t \left[ {\begin{array}{c}
1.6\\
2.5\\
1\\
\end{array} } \right]$
