---
title: "MThambeliyagodage_Data605_W2_Assign2"
author: "Matheesha Thambeliyagodage"
date: "February 10, 2018"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##  <span style="color:Green"> Problem set 1 </span>

### <span style="color:Green">Part (1)   </span>

#### <span style="color:blue"> Show That $\begin{aligned} A^tA \neq  AA^t \end{aligned}$ in general. (Proof and demonstration.) </span>
**Demonstration:**

$$ \begin{aligned}
      \text{Lets Consider the Matrix A = } 
        \begin{bmatrix} 1 & 0  & 3 & \\2  & -4  & 2\\1 & -3  & 2\end{bmatrix} \text{We know that } 
        A^T = \begin{bmatrix} 1 & 2  & 1 & \\0  & -4  & -3\\3 & 2  & 2\end{bmatrix} 
    \end{aligned}
$$

```{r}
A <- matrix(c(1,0,3,2,-4,2,1,-3,2), nrow=3, byrow=TRUE)
At <- t(A)
A
At
```
```{r}
# Multiplying A by its transpose we get 3x3
A %*% At

```
```{r}
# Multiplying transpose At by A we get 3x3
At %*% A 

```

\begin{aligned}
      \text{Then} 
        A^TA = \begin{bmatrix} 10 & 8  & 7 \\8  & 24  & 18\\7 & 18  & 14\end{bmatrix} 
        \text{and} 
        AA^T = \begin{bmatrix} 6 & -11  & 9 \\-11  & 25  & -14\\9 & -14  & 17\end{bmatrix}
    \end{aligned}
<span style="color:	#DC143C"> From the above it is obvious that  $\begin{aligned} A^tA \neq  AA^t \end{aligned}$ </span>

**Prof:**

\begin{aligned}
      \text{Lets Consider the Matrix A with i rows and j columns where } \begin{aligned} i\neq j \end{aligned} =
\begin{bmatrix}  a_{1,1} & a_{1,2} &...... &a_{1,j} \\
                         a_{2,1} & a_{2,2} &...... & a_{2,j}\\
                         a_{3,1} & a_{3,2} &...... & a_{3,j}\\
                         . & . &...... & .\\
                         . & . &...... & .\\
                         a_{i,1} & a_{i,2} &...... & a_{i,j}\\
                           
\end{bmatrix}
\end{aligned}



\begin{aligned}
      \text{Transpose of Matrix A with j rows and i columns where  } \begin{aligned} j\neq i is A^T_{(ji)} \end{aligned} 
 = \begin{bmatrix}  a_{1,1} & a_{2,1} &...... &a_{i,1} \\
                a_{1,2} & a_{2,2} &...... & a_{i,2}\\
                a_{1,3} & a_{2,3} &...... & a_{i,3}\\
                . & . &...... & .\\
                . & . &...... & .\\
                a_{1,j} & a_{2,j} &...... & a_{j,i}\\
                           
\end{bmatrix}
\end{aligned}


$$
  \begin{aligned} A{(ij)}A^t{(ji)} \end{aligned}\text{will result the matrix} \begin{aligned} AA^T{(ij)} \end{aligned}
$$
$$
\begin{aligned} A^t{(ji)}  A{(ij)} \end{aligned}\text{will result the matrix} \begin{aligned} A^TA{(ji)} \end{aligned}
$$
<span style="color:	#DC143C"> It is obvious that =>   $\begin{aligned} A{(ij)} A^t{(ji)} \neq A^t{(ji)}A{(ij)} \end{aligned}$ </span>

#### Example consider 4 x 5 matrix
```{r}
A = matrix(c(1,3,2,1, 1,2,1,1, 1,1,0,1, 1,2,2,1, 3,4,1,3), nrow=4, ncol=5)
A
```
```{r}
#transpose of matrix A, generates 5 X 4
tA = t(A)
tA
```
```{r}
#Create 4 X 4 matrix
A%*%tA
```
```{r}
#Create 5 X 5 matrix
tA%*%A
```
#### Outputs are different
 
### <span style="color:Green"> Part (2) </span>

#### <span style="color:blue">  For a special type of square matrix A, we get $\begin{aligned} A^tA =  AA^t \end{aligned}$ Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).
</span>

#### When the matrix is symmetrical square matrix $\begin{aligned} A^tA =  AA^t \end{aligned}$
**Prof:**
```{r}
# Symmetrical 4 X 4 matrix 
A = matrix(c(9,1,3,6, 1,11,7,6, 3,7,4,1, 6,6,1,10), nrow=4, ncol=4)
A
```
```{r}
#transpose of matrix A  is similar to A
tA = t(A)
tA
```
```{r}
#matrix multiplication
A%*%tA
```
```{r}
#matrix multiplication
tA%*%A
```
<span style="color:	#DC143C"> It is obvious that for Symmetrical matrix  =>   $\begin{aligned} A{(ii)} A^t{(ii)} = A^t{(ii)}A{(ii)} \end{aligned}$ </span>


#### When the matrix is identical $\begin{aligned} I^tI =  II^t \end{aligned}$
**Prof:**
```{r}
# Identical 4 X 4 matrix 
A = diag(4)
A
```
```{r}
# Identical 4 X 4 matrix 
At = t(A)
At
```
```{r}
#matrix multiplication
A%*%At
```
```{r}
#matrix multiplication
At%*%A
```

<span style="color:	#DC143C"> It is obvious that for Identical matrix  =>   $\begin{aligned} A{(ij)} A^t{(ji)} = A^t{(ji)}A{(ij)} \end{aligned}$ </span>

##  <span style="color:Green"> Problem set 2 </span>

### <span style="color:Green">Part (1)   </span>
Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track ights use a technique called Kalman Filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your Fight using radars.

**Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer.**
The function factorize checks for the size of a square matrix, and then factorizes it into LU.

```{r}
factorizeLU <- function(A) {
  # Check that A is square
  if (dim(A)[1]!=dim(A)[2]) {
    return(NA)
  }
  
  U <- A
  n <- dim(A)[1]
  L <- diag(n)
  
  # If dimension is 1, then U=A and L=[1]
  if (n==1) {
    return(list(L,U))
  }
  
  # Loop through the lower triangle (by rows and columns)
  # Determine multiplier for each position and add it to L
  for(i in 2:n) {
    for(j in 1:(i-1)) {
      multiplier <- -U[i,j] / U[j,j]
      U[i, ] <- multiplier * U[j, ] + U[i, ]
      L[i,j] <- -multiplier
    }
  }
  return(list(L,U))
}
```

### Demonstration

#### Example 1

```{r}
A <- matrix(c(1,4,-3,-2,8,5,3,4,7), nrow=3, byrow=TRUE)
LU <- factorizeLU(A)
L<-LU[[1]]  
U<-LU[[2]]

A
```
```{r}
L
```

```{r}
U
```
```{r}
L %*% U
```
```{r}
A == L %*% U
```

```{r}
A <- matrix(c(2,1,6,8),nrow=2, byrow = T)

LU <- factorizeLU(A)
L<-LU[[1]]  
U<-LU[[2]]
A
L
U
A == L %*% U
```





















