---
title: "DATA606_Meetup Presentation_6.11_Study_abroad"
author: "Matheesha Thambeliyagodage"
date: "May 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
#library(combinat)

normalPlot <- function(mean=0, sd=1, bounds=c(-1,1), tails=FALSE) {
	x <- seq(-4,4,length=100)*sd + mean
	hx <- dnorm(x,mean,sd)

	plot(x, hx, type="n", xlab="x-Axis", ylab="",
		 main="Normal Distribution", axes=FALSE)
	lines(x, hx)

	if(tails) {
		i.low <- x <= bounds[1]
		i.high <- x >= bounds[2]
		polygon(c(x[i.low],bounds[1]), c(hx[i.low], 0), col="red")
		polygon(c(bounds[2],x[i.high]), c(0,hx[i.high]), col="red")
	} else {
		i <- x >= bounds[1] & x <= bounds[2]
		polygon(c(bounds[1],x[i],bounds[2]), c(0,hx[i],0), col="red")
		area <- pnorm(bounds[2], mean, sd) - pnorm(bounds[1], mean, sd)
		if(diff(bounds) > 0) {
			result <- paste("P(",bounds[1],"< x <",bounds[2],") =",
							signif(area, digits=3))
			mtext(result,3)
		}
	}
	axis(1)
}

```



###6.11 Study abroad.
A survey on 1,509 high school seniors who took the SAT and who completed
an optional web survey between April 25 and April 30, 2007 shows that 55% of high school seniors
are fairly certain that they will participate in a study abroad program in college 

<b> (a) Is this sample a representative sample from the population of all high school seniors in the
US? Explain your reasoning.</b>   
  - Here Sample observations are independent. The sample doesnt represententire high school population. Not all high schoolers take the SAT; some take the ACT instead. It would be a good sample representation of all US high school seniors who took the SAT.
  
```{r}
n <-1509
p <- 0.55
success <- n*p
failure <- n*(1-p)
success
failure
```

  
<b> (b) Let's suppose the conditions for inference are met. Even if your answer to part (a) indicated
that this approach would not be reliable, this analysis may still be interesting to carry out
(though not report). Construct a 90% confidence interval for the proportion of high school
seniors (of those who took the SAT) who are fairly certain they will participate in a study
abroad program in college, and interpret this interval in context.</b>   

\(SE=\sqrt{\frac{p*(1-p)}{n}}\)

```{r}
se <-  sqrt((p*(1-p))/n)  ## Standard Error
z <- qnorm(.90)  ## Quantile Function - inverse of pnorm
CI.UPPER <- p+z*se
CI.LOWER<- p-z*se
se
z
CI.UPPER
CI.LOWER

CI <- c(CI.LOWER, CI.UPPER)
CI
```

<b> (c) What does "90% confidence" mean? </b>   

- It means that we are 90% confident that 53%-57% percent of high schools seniors are fairly like to participate in a study abroad program. We can say that 90% if random samples would give us 90% CI that includes the true proportion (.55).

```{r}

normalPlot(p,se,bounds = c(0.53,0.57),tails=FALSE) 
```

<b> (d) Based on this interval, would it be appropriate to claim that the majority of high school seniors
are fairly certain that they will participate in a study abroad program in college?</b>  
- Possible toclaim that the majority of HS seniors are fairly certain that they would participate in a study above program. What we found is the interval is over 50% (majority).
