---
title: "DATA606_W2_HW2_MatheeshaThambeliyagodage"
author: "Matheesha Thambeliyagodage"
date: "February 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**(Page-116)**

**2.2 Roulette wheel.** The game of roulette involves spinning a wheel with 38 slots: 18 red, 18
black, and 2 green. A ball is spun onto the wheel and will eventually land in a slot, where each
slot has an equal chance of capturing the ball.

Solution 1 > Each Spin is independent. There for P(getting a red slot ) is equal, that is 18/36
```{r}
18/36
```
Solution 2 > Even for 300 spins P(getting a red slot ) is 18/36
```{r}
18/36
```
Solution 3 > Not always, there could be unfair condition of the wheel or/and ball ..

**(Page-116)**

**2.3 Four games, one winner.** Below are four versions of the same game. Your archnemisis
gets to pick the version of the game, and then you get to choose how many times to flip a coin: 10
times or 100 times. Identify how many coin flips you should choose for each version of the game.
It costs $1 to play each game. Explain your reasoning.


**(a)** If the proportion of heads is larger than 0.60, you win $1.


**(b)** If the proportion of heads is larger than 0.40, you win $1.


**(c)** If the proportion of heads is between 0.40 and 0.60, you win $1.


**(d)** If the proportion of heads is smaller than 0.30, you win $1.


**Page 124**
** 2.38 Baggage fees.** An airline charges the following baggage fees: $25 for the first bag and
$35 for the second. Suppose 54% of passengers have no checked luggage, 34% have one piece of
checked luggage and 12% have two pieces. We suppose a negligible portion of people check more
than two bags.

(a) Build a probability model, compute the average revenue per passenger, and compute the
corresponding standard deviation.
```{r}
bags <- c(0, 1, 2)
prob <- c(0.54, 0.34, 0.12)
charg <- c(0, 25, 25 + 35)

probmod <- data.frame(bags, prob, charg)
probmod

# Compute the average revenue per passenger
probmod$avgrev <- probmod$prob * probmod$charg
probmod

sum(probmod$avgrev)


```

```{r}
#Compute Variance
probmod$mean <- probmod$weightRev - avgRevPerPax
probmod$DiffMeanSqrd <- probmod$DiffMean ^ 2
probmod$DiffMeanSqrdTimesProb <- probmod$DiffMeanSqrd * probmod$prob
probmod
# Compute standard deviation
varRevPerPax <- sum(probmod$DiffMeanSqrdTimesProb)
sdRevPerPax <- sqrt(varRevPerPax)
sdRevPerPax
```

```{r}
prob <- c(0.54, 0.34, 0.12)
bags <- c(0, 1, 2)
fees <- c(0, 25, 25 + 35)
df38 <- data.frame(prob, bags, fees)
df38$weightRev <- df38$prob * df38$fees
#open in browser PRO version Are you a developer? Try out the HTML to PDF API pdfcrowd.com
df38
# Compute the average revenue per passenger
avgRevPerPax <- sum(df38$weightRev)
avgRevPerPax
# Compute Variance
df38$DiffMean <- df38$weightRev - avgRevPerPax
df38$DiffMeanSqrd <- df38$DiffMean ^ 2
df38$DiffMeanSqrdTimesProb <- df38$DiffMeanSqrd * df38$prob
df38
# Compute standard deviation
varRevPerPax <- sum(df38$DiffMeanSqrdTimesProb)
sdRevPerPax <- sqrt(varRevPerPax)
sdRevPerPax

```


2.44 Income and gender. The relative frequency table below displays the distribution of
annual total personal income (in 2009 inflation-adjusted dollars) for a representative sample of
96,420,486 Americans. These data come from the American Community Survey for 2005-2009.
This sample is comprised of 59% males and 41% females.69
(a) Describe the distribution of total personal income.
(b) What is the probability that a randomly chosen US
resident makes less than $50,000 per year?
(c) What is the probability that a randomly chosen US
resident makes less than $50,000 per year and is female?
Note any assumptions you make.
(d) The same data source indicates that 71.8% of females
make less than $50,000 per year. Use this value to
determine whether or not the assumption you made in
part (c) is valid.



2.6 Dice rolls. If you roll a pair of fair dice, what is the probability of
(a) getting a sum of 1?
(b) getting a sum of 5?
(c) getting a sum of 12?


2.8 Poverty and language. The American Community Survey is an ongoing survey that
provides data every year to give communities the current information they need to plan investments
and services. The 2010 American Community Survey estimates that 14.6% of Americans live below
the poverty line, 20.7% speak a language other than English (foreign language) at home, and 4.2%
fall into both categories.59
(a) Are living below the poverty line and speaking a foreign language at home disjoint?
(b) Draw a Venn diagram summarizing the variables and their associated probabilities.
(c) What percent of Americans live below the poverty line and only speak English at home?
(d) What percent of Americans live below the poverty line or speak a foreign language at home?
(e) What percent of Americans live above the poverty line and only speak English at home?
(f) Is the event that someone lives below the poverty line independent of the event that the person
speaks a foreign language at home?


