---
title: "DATA606_W4_LAB4b_MatheeshaThambeliyagodage"
author: "Matheesha Thambeliyagodage"
date: "May 1, 2017"
output: html_document
---


```{r, echo=FALSE, message=FALSE} 
library(ggplot2)
myTheme <- theme(axis.ticks=element_blank(),  
                  panel.border = element_rect(color="gray", fill=NA), 
                  panel.background=element_rect(fill="#FBFBFB"), 
                  panel.grid.major.y=element_line(color="white", size=0.5), 
                  panel.grid.major.x=element_line(color="white", size=0.5))
set.seed(20275)

```

## The data

```{r load-data, eval=TRUE}
download.file("http://www.openintro.org/stat/data/ames.RData", destfile = "ames.RData")
load("ames.RData")

```


```{r}
n <- 60
population <- ames$Gr.Liv.Area
samp <- sample(population, 60)
summary(samp)

```

```{r}
hist(population)
hist(samp, breaks = 10)
```




```{r}
popMean <- mean(population)
popMean
```

```{r}
sample_mean <- mean(samp)
se <- sd(samp) / sqrt(60)
lower <- sample_mean - 1.96 * se
upper <- sample_mean + 1.96 * se
c(lower, upper)
```
```{r}
mean(population)
```

```{r}
samp_mean <- rep(NA, 50)
samp_sd <- rep(NA, 50)
n <- 60
```

Now we're ready for the loop where we calculate the means and standard deviations of 50 random samples.

```{r}
for(i in 1:50){
  samp <- sample(population, n) # obtain a sample of size n = 60 from the population
  samp_mean[i] <- mean(samp)    # save sample mean in ith element of samp_mean
  samp_sd[i] <- sd(samp)        # save sample sd in ith element of samp_sd
}
```

```{r}
lower_vector <- samp_mean - 1.96 * samp_sd / sqrt(n) 
upper_vector <- samp_mean + 1.96 * samp_sd / sqrt(n)
```

```{r}
c(lower_vector[1], upper_vector[1])
```

## On your own

1.Using the following function (which was downloaded with the data set), 
plot all intervals. What proportion of your confidence intervals include 
the true population mean? Is this proportion exactly equal to the 
confidence level? If not, explain why.
    
    
    
    
    
    
```{r}
plot_ci(lower_vector, upper_vector, mean(population))
```


2. Pick a confidence level of your choosing, provided it is not 95%. What is 
    the appropriate critical value?
    
    **Let's pick 95%. For this, the critical value will be 2.58.**
    
3. Calculate 50 confidence intervals at the confidence level you chose in the 
    previous question. You do not need to obtain new samples, simply calculate 
    new intervals based on the sample means and standard deviations you have 
    already collected. Using the `plot_ci` function, plot all intervals and 
    calculate the proportion of intervals that include the true population 
    mean. How does this percentage compare to the confidence level selected for
    the intervals?
    
    
```{r}
lower_vector95 <- samp_mean - 2.58 * samp_sd / sqrt(n) 
upper_vector95 <- samp_mean + 2.58 * samp_sd / sqrt(n)

```
```{r}
plot_ci(lower_vector95, upper_vector95, mean(population))
```
```{r}
capMean <- lower_vector95 < popMean && popMean < upper_vector95
sum(capMean) / length(capMean)
```
In this case all of the intervals captured the true mean, which is just 1% off from the confidence interval.


