---
title: "DATA607_W5_Assignment_Tidying_and_Transforming_Data"
author: "Matheesha Thambeliyagodage"
date: "March 25, 2017"
output: html_document
---


```{r}
# install.packages("stringr")
# install.packages("prettydoc")
# install.packages("tidyr")
# install.packages("dplyr")
# install.packages("knitr")


library(stringr)
library(prettydoc)
library(knitr)
library(tidyr)
library(dplyr)
```

#### 1). Uploaded .csv file from Github....
```{r}
my_data <- read.csv("https://raw.githubusercontent.com/mathsanu/CUNY_MSDA/master/DATA607/W5/flights.csv" ,stringsAsFactors = F)
my_data
```
#### 2). Display Data/Headings....
```{r}
str(my_data) 
```

#### 3). Assign two values to the missing cell values ....
```{r}
my_data[2, 1] <- "ALASKA"
my_data[5, 1] <- "AM WEST"
my_data
```

#### 4). Assign two new column names....

```{r}
names(my_data)[1:2] <- c('Airlines','Status')
my_data
```
#### 5). Gather flight details and cities in to two vectors....
```{r}
Tidydata<-gather(my_data, "cities","flights",3:7)
Tidydata
```

#### 6). Filter the rows with data into new Data Farme....
```{r}
Tidydata1 <- filter(Tidydata, Status !='')
Tidydata1
```

#### 7). With spread function Pivot Status column for the data in flight column and send the data to a new data frame....
```{r}
Tidydata2<-spread(Tidydata1, Status, flights)
Tidydata2
```
#### 8). Display data in two data frames in string format....
```{r}
glimpse(Tidydata1)
glimpse(Tidydata2)
```
#### 9). Display Tidydata1 - data in two data frames in table format....
```{r}
kable(Tidydata1)
```

#### 10). Select only given columns....
```{r}
select(Tidydata1,Airlines, cities)
filter(Tidydata2, cities=='Phoenix')
```
#### 11). Select distinct columns....
```{r}
distinct(Tidydata1,Airlines, cities)
```
#### 12). Calculate the ratio between Delayed and On_Time flights....
```{r}
Tidydata2 %>% 
mutate(ratio=delayed/(delayed+`on_time`)) %>%
  arrange(cities, Airlines)
```
#### 13). Plotting a graph....
```{r}
library(ggplot2)
Tidydata2 <- mutate(Tidydata2, Total = delayed + `on_time`, PercentDelayed = delayed / Total * 100)
Tidydata2 <- arrange(Tidydata2, cities, PercentDelayed)

ggplot(Tidydata2,aes(x=cities,y=PercentDelayed,fill=factor(Airlines)))+
    geom_bar(stat="identity",position="dodge")
```

#### 14).Export Data in to a .csv....

```{r}
# write.csv(my_data, 'E:/0_MSC/Modules/607- Data Acquisition and Management/Asignments/W5/FlightStatusTable.csv', row.names=T)

write.csv(my_data, "../W5/FlightStatusTable.csv")

```

#### 15). Group by airline and summarize the total on time flights divided by the total flights.....
```{r}
delays <- Tidydata2 %>% group_by(Airlines) %>% summarise(MeanPercent = round(mean(PercentDelayed), 0))
delays
```

