---
title: "MThambeliyagodage_Data605_W12_Assign12"
author: "Matheesha Thambeliyagodage"
date: "April 28, 2018"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
  word_document: 
    toc: yes
  pdf_document: 
    toc: yes
  html_notebook: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE)
options("scipen" = 10)
```
```{r}
library(ggplot2)
library(dplyr)
# library(gridExtra)
```
# Data 605 Assignment 12

### Task

Using the `cars` dataset in R, build a linear model for stopping distance as a function of speed and
replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and
residual analysis).

### Data Import

```{r}
# Import data
who <- read.csv('https://raw.githubusercontent.com/mathsanu/CUNY_MSDA/master/DATA605/W12/who.csv')

knitr::kable(head(who[,c(1,2,6,8,9,10)]))
```

Data is real-world World Health Organization data from 2008. It includes `r nrow(who)` observations for `r ncol(who)` variables. Data dictionary:

- `Country`: name of the country
- `LifeExp`: average life expectancy for the country in years
- `InfantSurvival`: proportion of those surviving to one year or more
- `Under5Survival`: proportion of those surviving to five years or more
- `TBFree`: proportion of the population without TB
- `PropMD`: proportion of the population who are MDs
- `PropRN`: proportion of the population who are RNs
- `PersExp`: mean personal expenditures on healthcare in US dollars at average exchange rate
- `GovtExp`: mean government expenditures per capita on healthcare, US dollars at average exchange rate
- `TotExp`: sum of personal and government expenditures

### Data Exploration

```{r}
summary(who)
```

Looking at the range of personal and government expenditures (13 to 482,750), I thought it was interesting to see top and bottom countries. 

```{r, echo=FALSE}
knitr::kable(head(who[order(who$TotExp),c(1,2,6,8,9,10)],5), caption="Bottom 5 Countries by Total Expenditures", row.names=FALSE)
knitr::kable(tail(who[order(who$TotExp),c(1,2,6,8,9,10)],5), caption="Top 5 Countries by Total Expenditures", row.names=FALSE)
```

### Question 1

Let us build a linear regression model for predicting life expectancy by total expenditures. Below scatterplot shows the relationship along with the linear regression line.

```{r}
# Linear regression model build
life_exp_lm <- lm(LifeExp ~ TotExp, data=who)

# Scatterplot of dependent and independent variables
plot(LifeExp~TotExp, data=who, 

     xlab="Total Expenditures", ylab="Life Expectancy",
     main="Life Expectancy vs Total Expenditures")
abline(life_exp_lm)

# Linear regression model summary
summary(life_exp_lm)

# Residuals variability plot
plot(life_exp_lm$fitted.values, life_exp_lm$residuals, 
     xlab="Fitted Values", ylab="Residuals",
     main="Residuals Plot")
abline(h=0)

# Residuals Q-Q plot
qqnorm(life_exp_lm$residuals)
qqline(life_exp_lm$residuals)
```

#### Results

**Residual standard error** is 9.371 and **F-statistic** is 65.26. Considering that average life expectancy is 67.38, the SE is not terrible and F-statistics is high. However, $R^2$ is only 0.2577 (so the model explains only 25.77% of variability). **P-value** is nearly 0, so the relationship is not due to random variation. 

Looking at residuals plots it is clear that there is no constant variability and that residuals are not normally distributed. This is **not a good model** to describe the relationship. It is clear from the scatterplot that the relationship is not linear.

### Question 2

Let us transform variables and re-run the simple linear regression model - $LifeExp^{4.6}$ and $TotExp^{0.06}$.

```{r}
# Transformation
LifeExp4.6 <- who$LifeExp^4.6
TotExp0.06 <- who$TotExp^0.06

# Linear regression model build
life_exp_lm <- lm(LifeExp4.6 ~ TotExp0.06)

# Scatterplot of dependent and independent variables
plot(LifeExp4.6~TotExp0.06, 
     xlab="Total Expenditures", ylab="Life Expectancy",
     main="Life Expectancy vs Total Expenditures (Transformed)")
abline(life_exp_lm)

# Linear regression model summary
summary(life_exp_lm)

# Residuals variability plot
plot(life_exp_lm$fitted.values, life_exp_lm$residuals, 
     xlab="Fitted Values", ylab="Residuals",
     main="Residuals Plot")
abline(h=0)

# Residuals Q-Q plot
qqnorm(life_exp_lm$residuals)
qqline(life_exp_lm$residuals)
```

#### Results

**Residual standard error** is 90,490,000 and **F-statistic** is 507.7. The F-statistic is good, but the SE is a bit high considering that it corresponds to 53.67 years if we reverse the transformation). $R^2$ is 0.7298, which is considerably better than in the first model (the model explains 72.98% of variability). **P-value** is again nearly 0, so the relationship is not due to random variation. 

Looking at residuals plots, variability is fairly constant with a few outliers and distribution of residuals is nearly normal with some deviation at the tails. This is **a fairly good model** to describe the relationship and it is significantly better than the first model. The linear relationship between transformed variables is clear from the scatterplot.


### Question 3

```{r}
newdata <- data.frame(TotExp0.06=c(1.5,2.5))
predict(life_exp_lm, newdata,interval="predict")^(1/4.6)
```

Based on the second model, prediction for total expeditures of \$860.705 ($TotExp^{0.06}=1.5$) is 63.31 years with a 95% confidence interval between 35.94 and 73.01.

Prediction for total expeditures of \$4,288,777 ($TotExp^{0.06}=2.5$) is 86.51 years with a 95% confidence interval between 81.81 and 90.43.


### Question 4

Let us build the following model: $LifeExp = \beta_0+\beta_1 \times PropMD + \beta_2 \times TotExp + \beta_3 \times PropMD \times TotExp$.

```{r}
# Multiple linear regression model build
life_exp_lm <- lm(LifeExp ~ PropMD + TotExp + TotExp:PropMD, data=who)

# Linear regression model summary
summary(life_exp_lm)

# Residuals variability plot
plot(life_exp_lm$fitted.values, life_exp_lm$residuals, 
     xlab="Fitted Values", ylab="Residuals",
     main="Residuals Plot")
abline(h=0)

# Residuals Q-Q plot
qqnorm(life_exp_lm$residuals)
qqline(life_exp_lm$residuals)
```

#### Results 

**Residual standard error** is 8.765 and **F-statistic** is 34.49. Considering that average life expectancy is 67.38, the SE is not terrible and F-statistics is fairly high (but lower than in the first model). $R^2$ is only 0.3574, so the model explains only 35.74% of variability, which is not high. **P-value** is nearly 0, so the relationship is not due to random variation. 

Looking at residuals plots it is clear that there is no constant variability and that residuals are not normally distributed. This is **not a good model** to describe the relationship. Kind of similar to the first model.


### Question 5

Consider forecast based on the last model with $PropMD = 0.03$ and $TotExp = 14$.

```{r}
newdata <- data.frame(PropMD=0.03, TotExp=14)
predict(life_exp_lm, newdata,interval="predict")
```

The prediction is 107.70 years with 95% confidence interval between 84.25 and 131.14. The prediction is completely unrealistic. We do have individuals livings into their 100s; however, consider that the total expenditures of \$14 is just a tad higher than the minimum value of \$13 for Burundi and the life expectancy there is 49 years. The highest life expectancy in the data is 83 years. There is nothing in our data to support this prediction and it goes against common sense. As stated under question 4, this is not a good model.