---
title: "DATA606_W2_Lab2_MatheeshaThambeliyagodage"
author: "Matheesha Thambeliyagodage"
date: "February 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
```{r}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")
head(kobe)
summary(kobe)
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak))
```


## On your own
 Comparing Kobe Bryant to the Independent Shooter Using calc_streak, compute the streak lengths of sim_basket.  
 
####  (1) Describe the distribution of streak lengths. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? How long is the player's longest streak of baskets in 133 shots?
```{r}
#outcomes <- c("H", "M")
#outcomes
#sample(outcomes, size = 1, replace = TRUE)
#sim_basket <- sample(outcomes, size = 130, replace = TRUE)
#sim_basket

outcomes <- c("H", "M")
outcomes
sim_basket <- sample(outcomes, size = 130, replace=T, prob = c(0.45, 0.55))
sim_basket

```



```{r}
calc_streak(sim_basket)
```

```{r}
sim_streak <- calc_streak(sim_basket)
barplot(table(sim_streak))
```


* The distribution is right-skewed. The typical streak length for this kind of shooter is 0 and 1, as they are the most common/typical values. The longest streak was 6 baskets.

####  (2) If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.

* There could be some changes in values of the sample, which is simulated randomley. But the final results may not have huge variation to the above first Experiment. If we set different values to the shooting percentage then we could see notisable variation


####  (2) How does Kobe Bryant's distribution of streak lengths compare to the distribution of streak lengths for the simulated shooter? Using this comparison, do you have evidence that the hot hand model fits Kobe's shooting patterns? Explain.

* Here now I am comparing two baskets kobe_streak  and sim_streak
```{r}
kobe_streak = calc_streak(kobe$basket)
sim_streak = calc_streak(sim_basket)
# Compute summaries:
summary(kobe_streak)
summary(sim_streak)
```
```{r}
# Make bar plots:
kobe_table = table(kobe_streak)
sim_table = table(sim_streak)

barplot(kobe_table)


```

```{r}
barplot(sim_table)
```

* Since Kobe's streak length distribution looks very similar to the Sim's streak  simulated steak length distribution, we can conclude that Kobe Bryant likely does not have a "hot hand".

